Solving an Exam Scheduling Problem Using a Genetic Algorithm

### 2009-12-02 trunk/Paper/paper.tex
Paper update

      The neighbouring schedules are those where only one course meets in either a different
      room, on a different day, or at a different time. 
%      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
      It can be seen that this method is completely deterministic given a starting schedule,
      and will always return a local maximum.

@@

      This method can be changed to true simulated annealing by varying the degree of
      modification made to the current best schedule with the number of evaluations
      remaining. Not being the focus of our investigation, little effort was made to 
      optimize the parameters. 1\% does seem to give better performance than other nearby 
      values, though we did not collect data to substantiate this. Finding ideal parameters 
      for simulated annealing on this problem would make for a project-sized inquiry in itself.

@@

    The sized of our 3 instances' vary from manageable to large. The total number 
    of different possible schedules in any problem instance is $|C|^{|R|dt}$ which 
    is far too large to conduct a brute force search. The smallest could potentially 
    be solved exactly, with methods not under consideration here, while the others 
    have search spaces much too large for any hope of finding a perfect solution. 
    For the largest, which simulates a UofT exam schedule, a single evaluation of 
    the fitness function takes 0.2 seconds on on a typical desktop computer. This 
    makes it difficult to thoroughly investigate this instance.

    The large data set was created with parameters corresponding to the fall 2009 
	University of Toronto exam schedule, which is the size we expect for a large 
	real-world problem instance.

    \definecolor{light-gray}{gray}{0.87}
    \rowcolors{1}{light-gray}{white}
    \begin{table}[h]
    \begin{small}
    \begin{tabular}{*{7}{c}}
    \hline 
 	  	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
    \hline \hline
     d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
     t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
     $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
     $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
     $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
     $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
     \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
     \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
    \hline  
    \end{tabular} 
	\end{small}
	\caption{The six data sets used in our project. $\#$Maj is the number of different 
	         majors used and \textsc{SSS} is an approximation of the search space size.}
	\end{table}

@@

    \begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
      \caption{Comparing the four searches on a small data set}
	\end{figure}

	\begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
      \caption{Comparing the four searches on a large data set}
	\end{figure}

>>>>

      The neighbouring schedules are those where only one course meets in either a different
      room, on a different day, or at a different time. 
      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.

      It can be seen that this method is completely deterministic given a starting schedule,
      and will always return a local maximum.

@@

      This method can be changed to true simulated annealing by varying the degree of
      modification made to the current best schedule with the number of evaluations
      remaining. Not being the focus of our investigation, little effort was made to 
      optimize the parameters. 
      %1\% does seem to give better performance than other nearby 
%      values, though we did not collect data to substantiate this. Finding ideal parameters 
 %     for simulated annealing on this problem would make for a project-sized inquiry in itself.

@@

    The size of our 3 instances vary considerably. The total number of different possible
    schedules in any problem instance is $|C|^{|R|dt}$ which is far too large to conduct
    a brute force search. The smallest could potentially be solved exactly, with methods
    not under consideration here, while the others have search spaces much too large 
    for any hope of finding a perfect solution. For the largest, which simulates a UofT 
    exam schedule, a single evaluation of the fitness function takes 0.2 seconds on on 
    a typical desktop computer. This makes it difficult to thoroughly investigate this 
    instance.

    The large problem instance was created with parameters corresponding to the fall 2009 
	University of Toronto exam schedule, which is the size we expect for a large 
	real-world problem instance.

    \definecolor{light-gray}{gray}{0.87}
    \rowcolors{1}{light-gray}{white}
    \begin{table}[h]
    \begin{small}
    \begin{tabular}{*{7}{c}}
    \hline 
 	  	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
    \hline \hline
     d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
     t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
     $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
     $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
     $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
     $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
     \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
     \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
    \hline  
    \end{tabular} 
	\end{small}
	\caption{The six problem instances used in our project. $\#$Maj is the number of different 
	         majors used and \textsc{SSS} is an approximation of the search space size.}
	\end{table}

@@

    \begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
      \caption{Search methods on small problem instance}
	\end{figure}

	\begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
      \caption{Search methods on large problem instance}
	\end{figure}
